
```{r eval=FALSE}

# conda activate voyager
library(Seurat)
library(tidyverse)
library(RColorBrewer)
library(viridis)
library(cowplot)
library(ggrepel)
library(ggpubr)
library(ggrastr)
library(patchwork)
theme_set(theme_cowplot())

# set random seed for reproducibility
set.seed(12345)

setwd('/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/')
fig_dir <- 'figures/'
data_dir <- 'data/'

# load seurat object
seurat_obj <- readRDS(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/data/harmony_annotated_integration.rds')

# load the color scheme!!
load('/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/data/color_scheme.rda')

```

Combine the individual DEG tables

```{r eval=FALSE}

###############################################################################
# Cluster markers 
###############################################################################

DEG_dir <- "/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/cluster_markers/"
DEG_tests <- dir(DEG_dir)

combined <- Reduce(rbind, lapply(DEG_tests, function(file){
  read.csv(paste0(DEG_dir, '/', file))
}))
combined$Nr4a2 <- ifelse(combined$gene %in% Nr4a2_targets, TRUE, FALSE)

write.csv(
  combined,
  file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_marker_DEGs.csv',
  quote=FALSE, row.names=FALSE
)


###############################################################################
# celltype markers 
###############################################################################

DEG_dir <- "/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/celltype_markers/"
DEG_tests <- dir(DEG_dir)

combined <- Reduce(rbind, lapply(DEG_tests, function(file){
  read.csv(paste0(DEG_dir, '/', file))
}))
combined$Nr4a2 <- ifelse(combined$gene %in% Nr4a2_targets, TRUE, FALSE)

write.csv(
  combined,
  file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_marker_DEGs.csv',
  quote=FALSE, row.names=FALSE
)

###############################################################################
# Cluster Nurr2c vs GFP
###############################################################################

DEG_dir <- "/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/cluster_Nurr2c_vs_GFP/"
DEG_tests <- dir(DEG_dir)

combined <- Reduce(rbind, lapply(DEG_tests, function(file){
  read.csv(paste0(DEG_dir, '/', file))
}))
combined$Nr4a2 <- ifelse(combined$gene %in% Nr4a2_targets, TRUE, FALSE)

write.csv(
  combined,
  file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Nurr2c_vs_GFP_DEGs.csv',
  quote=FALSE, row.names=FALSE
)

###############################################################################
# Cluster Naive Nurr2c vs GFP
###############################################################################

DEG_dir <- "/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/cluster_Naive_Nurr2c_vs_GFP/"
DEG_tests <- dir(DEG_dir)

combined <- Reduce(rbind, lapply(DEG_tests, function(file){
  read.csv(paste0(DEG_dir, '/', file))
}))
combined$Nr4a2 <- ifelse(combined$gene %in% Nr4a2_targets, TRUE, FALSE)

write.csv(
  combined,
  file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Naive_Nurr2c_vs_GFP_DEGs.csv',
  quote=FALSE, row.names=FALSE
)


###############################################################################
# Celltype Nurr2c vs GFP
###############################################################################

DEG_dir <- "/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/celltype_Nurr2c_vs_GFP/"
DEG_tests <- dir(DEG_dir)

combined <- Reduce(rbind, lapply(DEG_tests, function(file){
  read.csv(paste0(DEG_dir, '/', file))
}))
combined$Nr4a2 <- ifelse(combined$gene %in% Nr4a2_targets, TRUE, FALSE)

write.csv(
  combined,
  file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv',
  quote=FALSE, row.names=FALSE
)

###############################################################################
# Celltype Naive Nurr2c vs GFP
###############################################################################

DEG_dir <- "/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/celltype_Naive_Nurr2c_vs_GFP/"
DEG_tests <- dir(DEG_dir)

combined <- Reduce(rbind, lapply(DEG_tests, function(file){
  read.csv(paste0(DEG_dir, '/', file))
}))
combined$Nr4a2 <- ifelse(combined$gene %in% Nr4a2_targets, TRUE, FALSE)

write.csv(
  combined,
  file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv',
  quote=FALSE, row.names=FALSE
)

```

Write the significant DEGs for the Supplementary Tables

```{r eval=FALSE}

# cluster DEGs
degs_cl <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Nurr2c_vs_GFP_DEGs.csv')
degs_ct <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv')

keep_groups <- setdiff(unique(degs_cl$group), unique(degs_ct$group))
degs_cl <- subset(degs_cl, p_val_adj < 0.05 & group %in% keep_groups) %>% 
  dplyr::select(-Nr4a2)
degs_ct <- subset(degs_ct, p_val_adj < 0.05) %>% 
    dplyr::select(-Nr4a2)
degs <- rbind(degs_cl, degs_ct)

write.csv(
  degs,
  file=paste0(data_dir, 'experienced_DEGs_signif.csv'),
  quote=FALSE, row.names=FALSE
)

# celltype DEGs

# Naive cluster DEGs
degs_cl <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs_ct <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv')

keep_groups <- setdiff(unique(degs_cl$group), unique(degs_ct$group))
degs_cl <- subset(degs_cl, p_val_adj < 0.05 & group %in% keep_groups) %>% 
  dplyr::select(-Nr4a2)
degs_ct <- subset(degs_ct, p_val_adj < 0.05) %>% 
    dplyr::select(-Nr4a2)
degs <- rbind(degs_cl, degs_ct)

write.csv(
  degs,
  file=paste0(data_dir, 'naive_DEGs_signif.csv'),
  quote=FALSE, row.names=FALSE
)

```

Plot the Marker gene heatmap (Figure 2B)

```{r eval=FALSE}

# read combined
degs <- read.csv('/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_marker_DEGs.csv')
group_levels <- levels(seurat_obj$cell_identity)
degs$group <- factor(as.character(degs$group), levels=group_levels)

n_degs <- 5
plot_genes <- degs %>%
  arrange(group) %>%
  subset(p_val_adj <= 0.05) %>%
  group_by(group) %>%
  top_n(n_degs, wt=avg_log2FC)  %>%
  .$gene


seurat_obj$barcode <- colnames(seurat_obj)
temp <- table(seurat_obj@meta.data$cell_identity)

df <- data.frame()
for(i in 1:length(temp)){

  if(temp[[i]] < 1000){
    cur_df <- seurat_obj@meta.data %>% subset(cell_identity == names(temp)[i])
  } else{
    cur_df <- seurat_obj@meta.data %>% subset(cell_identity == names(temp)[i]) %>% sample_n(1000)
  }
  df <- rbind(df, cur_df)
}

seurat_obj <- ScaleData(seurat_obj, features=plot_genes)

p <- DoHeatmap(
  seurat_obj %>% subset(barcode %in% df$barcode),
  features=unlist(plot_genes),
  group.by='cell_identity',
  group.colors = cell_group_colors,
  raster=TRUE, slot='scale.data'
) + theme(
  axis.text.y = element_text(face='italic', size=3)
) 

pdf(paste0(fig_dir, 'cluster_marker_gene_heatmap.pdf'), width=14, height=8, useDingbats=FALSE)
p
dev.off()

```

Neuronal marker gene analysis (Figure S2)

```{r eval=FALSE}

#--------------------------------------------------------------------------------
# MHb markers
#--------------------------------------------------------------------------------

seurat_mhb <- subset(seurat_obj, cell_type == 'MHb-Neuron')
seurat_mhb$annotation <- droplevels(seurat_mhb$annotation)
Idents(seurat_mhb) <- seurat_mhb$annotation

markers <- FindAllMarkers(
  seurat_mhb,
  test.use = "MAST",
  min.pct = 0.2,
  logfc.threshold = 0.5,
  only.pos =TRUE
)

write.csv(markers, quote=FALSE, file=paste0(data_dir, 'mhb_markers.csv'))
markers <- read.csv(file=paste0(data_dir, 'mhb_markers.csv'))

# exclude mitochondrial genes from the plot
markers <- markers[!grepl("^mt-", markers$gene),]
n_degs <- 15
plot_genes <- markers %>%
  arrange(cluster) %>%
  subset(p_val_adj <= 0.05) %>%
  group_by(cluster) %>%
  top_n(n_degs, wt=avg_log2FC)  %>%
  .$gene

# set random seed
set.seed(42)

seurat_mhb$barcode <- colnames(seurat_mhb)
temp <- seurat_mhb@meta.data %>% group_by(annotation) %>% sample_n(300)

seurat_mhb <- ScaleData(seurat_mhb, features=plot_genes)

seurat_mhb$ordered_clusters <- fct_rev(seurat_mhb$annotation)
p <- DoHeatmap(
  seurat_mhb %>% subset(barcode %in% temp$barcode),
  features=unlist(plot_genes),
  group.by='annotation',
  raster=TRUE, slot='scale.data',
  group.colors=cluster_colors
)+ theme(axis.text.y = element_text(face='italic'))

pdf(paste0(fig_dir, 'mhb_marker_gene_heatmap.pdf'), width=7, height=12, useDingbats=FALSE)
p
dev.off()

subset(markers, cluster == 'MHb-5') %>% arrange(avg_log2FC)

#--------------------------------------------------------------------------------
# LHb markers
#--------------------------------------------------------------------------------

seurat_lhb <- subset(seurat_obj, cell_type == 'LHb-Neuron')
seurat_lhb$annotation <- droplevels(seurat_lhb$annotation)
Idents(seurat_lhb) <- seurat_lhb$annotation

markers <- FindAllMarkers(
  seurat_lhb,
  test.use = "MAST",
  min.pct = 0.2,
  logfc.threshold = 0.5,
  only.pos =TRUE
)

write.csv(markers, quote=FALSE, file=paste0(data_dir, 'lhb_markers.csv'))
markers <- read.csv(file=paste0(data_dir, 'lhb_markers.csv'))


markers <- markers[!grepl("^mt-", markers$gene),]
n_degs <- 10
plot_genes <- markers %>%
  arrange(cluster) %>%
  subset(p_val_adj <= 0.05) %>%
  group_by(cluster) %>%
  top_n(n_degs, wt=avg_log2FC)  %>%
  .$gene

# set random seed
set.seed(42)

seurat_lhb$barcode <- colnames(seurat_lhb)
temp <- seurat_lhb@meta.data %>% group_by(annotation) %>% sample_n(300)
seurat_lhb <- ScaleData(seurat_lhb, features=plot_genes)

seurat_lhb$ordered_clusters <- fct_rev(seurat_lhb$annotation)
p <- DoHeatmap(
  seurat_lhb %>% subset(barcode %in% temp$barcode),
  features=unlist(plot_genes),
  group.by='annotation',
  raster=TRUE, slot='scale.data',
  group.colors=cluster_colors
)+ theme(axis.text.y = element_text(face='italic'))

pdf(paste0(fig_dir, 'lhb_marker_gene_heatmap.pdf'), width=7, height=12, useDingbats=FALSE)
p
dev.off()

subset(markers, cluster == 'MHb-5') %>% arrange(avg_log2FC)


#--------------------------------------------------------------------------------
# PHb markers
#--------------------------------------------------------------------------------

seurat_phb <- subset(seurat_obj, cell_type == 'PHb-Neuron')
seurat_phb$annotation <- droplevels(seurat_phb$annotation)
Idents(seurat_phb) <- seurat_phb$annotation

markers <- FindAllMarkers(
  seurat_phb,
  test.use = "MAST",
  min.pct = 0.2,
  logfc.threshold = 0.5,
  only.pos =TRUE
)

write.csv(markers, quote=FALSE, file=paste0(data_dir, 'phb_markers.csv'))
markers <- read.csv(file=paste0(data_dir, 'phb_markers.csv'))

markers <- markers[!grepl("^mt-", markers$gene),]

n_degs <- 15
plot_genes <- markers %>%
  arrange(cluster) %>%
  subset(p_val_adj <= 0.05) %>%
  group_by(cluster) %>%
  top_n(n_degs, wt=avg_log2FC)  %>%
  .$gene



# set random seed
set.seed(42)

seurat_phb$barcode <- colnames(seurat_phb)
temp <- seurat_phb@meta.data %>% group_by(annotation) %>% sample_n(300)
seurat_phb <- ScaleData(seurat_phb, features=plot_genes)

seurat_phb$ordered_clusters <- fct_rev(seurat_phb$annotation)
p <- DoHeatmap(
  seurat_phb %>% subset(barcode %in% temp$barcode),
  features=unlist(plot_genes),
  group.by='annotation',
  raster=TRUE, slot='scale.data',
  group.colors=cluster_colors
) + theme(axis.text.y = element_text(face='italic'))

pdf(paste0(fig_dir, 'phb_marker_gene_heatmap.pdf'), width=7, height=12, useDingbats=FALSE)
p
dev.off()


```

Make a table of primary and secondary Nr4a2 target genes using the 
TF network analysis.

```{r eval=FALSE}

cur_tf <- 'Nr4a2'

tf_nets <- dir('../tf_net/data/tf_nets/')
tf_nets <- tf_nets[grepl('importance', tf_nets)]

network_df <- data.frame()

for(cur_net_file in tf_nets){

  print(cur_net_file)

  tmp <- strsplit(cur_net_file, '_')[[1]]
  cur_celltype <- tmp[2]
  cur_group <- tmp[3]

  importance_df <- read.csv(paste0('../tf_net/data/tf_nets/', cur_net_file ))

  #---------------------------------------------------------------------------#
  # Define the TF regulons
  #---------------------------------------------------------------------------#

  n_tfs <- 5
  importance_thresh <- 0.001

  regulons <- importance_df %>% 
    subset(Gain > importance_thresh) %>% 
    group_by(gene) %>%
    slice_max(order_by=Gain, n=n_tfs) %>% 
    ungroup()

  # compute the degree for each TF:
  tf_degrees <- table(regulons$tf)

  #---------------------------------------------------------------------------#
  # Get the primary & secondary targets of Nr4a2
  #---------------------------------------------------------------------------#

  # primary target genes 
  cur_primary<- regulons %>% 
    subset(tf == cur_tf) 

  # which of these pimary target genes are tfs?
  cur_primary_tfs <- cur_primary %>% 
    subset(gene %in% unique(regulons$tf)) %>% .$gene

  cur_tfs <- unique(c(cur_tf, cur_primary_tfs))

  # get the regulons for these TFs:
  cur_secondary <- subset(regulons, tf %in% cur_primary_tfs)
  cur_secondary_tfs <- cur_primary %>% 
    subset(gene %in% unique(regulons$tf)) %>% .$gene

  # combine the primary and secondary into one table 
  cur_network <- rbind(cur_primary, cur_secondary)
  cur_network$Gain <- cur_network$Gain * sign(cur_network$Cor)

  cur_genes <- unique(cur_network$gene)
  length(cur_genes)

  # make an igraph network from the nr4a2 regulon:
  cur_network <- cur_network %>%
    dplyr::rename(c(source=tf, target=gene)) %>%
    mutate(Score = sign(Cor) * Gain)

  primary_genes <- unique(cur_primary$gene)
  secondary_genes <- unique(cur_network$target[! cur_network$target %in% primary_genes])

  cur_network$target_type <- ifelse(cur_network$target %in% primary_genes, 'primary', 'secondary')
  cur_network$cell_group <- cur_celltype 
  cur_network$group <- cur_group

  network_df <- rbind(network_df, cur_network)
}

# save this as a supp table!
df <- network_df %>% subset(group != 'PHb-Neuron')
df <- df %>%  dplyr::select(-c(Cover, Frequency, Cor, Score)) 
df <- df %>% dplyr::rename(regulatory_score=Gain)

write.csv(df, file=paste0(data_dir, 'Nr4a2_regulons.csv'), quote=FALSE, row.names=FALSE)

```

Nurr2c vs GFP with updated Nr4a2 targets (Neurons only)
Volcano plots in Figures 4 and 5

```{r eval=FALSE}

#---------------------------------------------------------------------------#
# First select one of the following groups of DEGs for plotting
#---------------------------------------------------------------------------#

# cluster DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_identity)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_identity)
)
name <- 'cluster'
Nr4a2 <- FALSE
nlabel <- 5
color1 <- 'darkgoldenrod3'; color2 <- 'hotpink3'

# celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype'
Nr4a2 <- FALSE
nlabel <- 5
color1 <- 'darkgoldenrod3'; color2 <- 'hotpink3'

# Naive cluster DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_identity)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_identity)
)
name <- 'cluster_naive'
Nr4a2 <- TRUE
nlabel <- 5
color2 <- "#643D78"; color1 <- "#14703F"

# Naive celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype_naive'
Nr4a2 <- TRUE
nlabel <- 5
color2 <- "#643D78"; color1 <- "#14703F"


# subset the neurons (this doesn't work for cluster level info)
degs <- subset(degs, group %in% c('MHb-Neuron', 'LHb-Neuron', 'PHb-Neuron'))
degs$group <- droplevels(degs$group)

#---------------------------------------------------------------------------#
# Add the TF information to the deg table
#---------------------------------------------------------------------------#

clusters <- unique(degs$group)

tmp <- data.frame()
for(cur_celltype in clusters){
  print(cur_celltype)
  group1 <- unique(degs$ident1)
  group2 <- unique(degs$ident2)

  # get the DEGs for this cluster
  cur_degs <- subset(degs, group == cur_celltype)

  # for this cluster, get the set of primary and secondary Nr4a2 targets:
  cur_network <- subset(network_df, cell_group == cur_celltype & group %in% c(group1, group2))
  primary_genes <- subset(cur_network, target_type == 'primary') %>% .$target
  secondary_genes <- subset(cur_network, target_type == 'secondary') %>% .$target

  # add info to the deg table:
  cur_degs$target_type <- ifelse(cur_degs$gene %in% secondary_genes, 'secondary', 'other')
  cur_degs$target_type <- ifelse(cur_degs$gene %in% primary_genes, 'primary', cur_degs$target_type)

  tmp <- rbind(tmp, cur_degs)

}
degs <- tmp

#---------------------------------------------------------------------------#
# subset by Nr4a2 targets?
#---------------------------------------------------------------------------#

cur_target_type <- 'primary'
cur_target_type <- 'secondary'
cur_target_type <- 'other'


degs <- subset(degs, target_type == cur_target_type)

#---------------------------------------------------------------------------#
# setup table for volcano plot
#---------------------------------------------------------------------------#

# remove the mt genes from the volcano plot
degs <- degs[!grepl('mt-', degs$gene),]

# lowest non-zero value
lowest <- degs %>% subset(p_val_adj != 0) %>% top_n(-1, wt=p_val_adj) %>% .$p_val_adj
degs$p_val_adj <- ifelse(degs$p_val_adj == 0, lowest, degs$p_val_adj)


# label the top and bottom significant genes by log fold change
cur_degs <- Reduce(rbind, lapply(unique(degs$group), function(x){
  cur <- subset(degs, group == x)

  top_thresh <- cur %>% subset(p_val_adj <= 0.05 & avg_log2FC > 0) %>% top_n(nlabel, wt=avg_log2FC) %>% .$avg_log2FC %>% min
  bottom_thresh <- cur %>% subset(p_val_adj <= 0.05 & avg_log2FC < 0) %>% top_n(-1*nlabel, wt=avg_log2FC) %>% .$avg_log2FC %>% max

  cur$anno <- ifelse(cur$p_val_adj <= 0.05 & cur$avg_log2FC >= top_thresh, cur$gene, NA)
  cur$anno <- ifelse(cur$p_val_adj <= 0.05 & cur$avg_log2FC <= bottom_thresh, cur$gene, cur$anno)
  cur$anno <- ifelse(cur$gene == 'Nr4a2', cur$gene, cur$anno)
  cur$color <- ifelse(cur$p_val_adj > 0.05, 'gray', ifelse(cur$avg_log2FC > 0, color1, color2))
  cur
}))

groups <- levels(degs$group)
plot_list <- list()
for(cluster  in groups){


  print(cluster)
  plot_degs <- cur_degs %>% subset(group == cluster)

  p <- plot_degs  %>%
     ggplot(aes(x=avg_log2FC, y=-log10(p_val_adj))) +
     geom_hline(yintercept=-log10(0.05), linetype='dashed')

  # plot genes that are Nr4a2 targets
  p <- p + ggrastr::rasterise(geom_point(
    alpha=0.5,
    color=plot_degs %>% .$color
  ), dpi=500)

  p <- p +
     geom_point(
       inherit.aes=FALSE,
       data=subset(plot_degs, !is.na(anno)),
       aes(avg_log2FC, -log10(p_val_adj)),
       fill=subset(plot_degs, !is.na(anno)) %>% .$color,
       shape=21, size=3, color='black'
     ) +
     geom_text_repel(aes(label=anno), color='black', fontface='italic',  min.segment.length=0) +
     xlim(-1*max(abs(plot_degs$avg_log2FC))-0.1, max(abs(plot_degs$avg_log2FC))+0.1) +
     ggtitle(paste0(cluster)) +
     xlab(bquote("Average log"[2]~"(Fold Change)")) +
     ylab(bquote("-log"[10]~"(Adj. P-value)")) +
     theme(
       panel.border = element_rect(color='black', fill=NA, size=1),
       panel.grid.major = element_blank(),
       axis.line = element_blank(),
       plot.title = element_text(hjust = 0.5),
       legend.position='bottom'
     )

    plot_list[[cluster]] <- p

}


 out <- paste0(fig_dir, 'volcano_', name, '_', cur_target_type, '.pdf') 

plot_list <- lapply(plot_list, function(x){
  x + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(0,0,0,0),
    plot.title = element_text(vjust=-0.2)
  )
})

# cluster
# pdf(out, width=18, height=12, useDingbats=FALSE)
# wrap_plots(plot_list, ncol=6)
# dev.off()

# celltype
pdf(out, width=9, height=3, useDingbats=FALSE)
wrap_plots(plot_list, ncol=3)
dev.off()

```

Volcano plots for all cell types, not split by TF target type
(Figure S6)

```{r eval=FALSE}

# cluster DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_identity)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_identity)
)
name <- 'cluster'
Nr4a2 <- FALSE
nlabel <- 5
color1 <- 'darkgoldenrod3'; color2 <- 'hotpink3'

# celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype'
Nr4a2 <- FALSE
nlabel <- 5
color1 <- 'darkgoldenrod3'; color2 <- 'hotpink3'

# Naive cluster DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_identity)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_identity)
)
name <- 'cluster_naive'
Nr4a2 <- TRUE
nlabel <- 5
color2 <- "#643D78"; color1 <- "#14703F"


# Naive celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype_naive'
Nr4a2 <- TRUE
nlabel <- 5
color2 <- "#643D78"; color1 <- "#14703F"


# remove the mt genes from the plot
degs <- degs[!grepl('mt-', degs$gene),]

# lowest non-zero value
lowest <- degs %>% subset(p_val_adj != 0) %>% top_n(-1, wt=p_val_adj) %>% .$p_val_adj
degs$p_val_adj <- ifelse(degs$p_val_adj == 0, lowest, degs$p_val_adj)

# label the top and bottom significant genes by log fold change
cur_degs <- Reduce(rbind, lapply(unique(degs$group), function(x){
  cur <- subset(degs, group == x)

  top_thresh <- cur %>% subset(p_val_adj <= 0.05 & avg_log2FC > 0) %>% top_n(nlabel, wt=avg_log2FC) %>% .$avg_log2FC %>% min
  bottom_thresh <- cur %>% subset(p_val_adj <= 0.05 & avg_log2FC < 0) %>% top_n(-1*nlabel, wt=avg_log2FC) %>% .$avg_log2FC %>% max

  cur$anno <- ifelse(cur$p_val_adj <= 0.05 & cur$avg_log2FC >= top_thresh, cur$gene, NA)
  cur$anno <- ifelse(cur$p_val_adj <= 0.05 & cur$avg_log2FC <= bottom_thresh, cur$gene, cur$anno)
  cur$anno <- ifelse(cur$gene == 'Nr4a2', cur$gene, cur$anno)
  cur$color <- ifelse(cur$p_val_adj > 0.05, 'gray', ifelse(cur$avg_log2FC > 0, color1, color2))
  cur
}))

groups <- levels(degs$group)
plot_list <- list()
for(cluster  in groups){


  print(cluster)
  plot_degs <- cur_degs %>% subset(group == cluster)

  p <- plot_degs  %>%
     ggplot(aes(x=avg_log2FC, y=-log10(p_val_adj))) +
     geom_hline(yintercept=-log10(0.05), linetype='dashed')

  # plot genes that are Nr4a2 targets
  p <- p + ggrastr::rasterise(geom_point(
    alpha=0.5,
    color=plot_degs %>% .$color
  ), dpi=500)

  p <- p +
     geom_point(
       inherit.aes=FALSE,
       data=subset(plot_degs, !is.na(anno)),
       aes(avg_log2FC, -log10(p_val_adj)),
       fill=subset(plot_degs, !is.na(anno)) %>% .$color,
       shape=21, size=3, color='black'
     ) +
     geom_text_repel(aes(label=anno), color='black', fontface='italic',  min.segment.length=0) +
     xlim(-1*max(abs(plot_degs$avg_log2FC))-0.1, max(abs(plot_degs$avg_log2FC))+0.1) +
     ggtitle(paste0(cluster)) +
     xlab(bquote("Average log"[2]~"(Fold Change)")) +
     ylab(bquote("-log"[10]~"(Adj. P-value)")) +
     theme(
       panel.border = element_rect(color='black', fill=NA, size=1),
       panel.grid.major = element_blank(),
       axis.line = element_blank(),
       plot.title = element_text(hjust = 0.5),
       legend.position='bottom'
     )

    plot_list[[cluster]] <- p

}

  out <- paste0(fig_dir, 'volcano_', name, '.pdf')



plot_list <- lapply(plot_list, function(x){
  x + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(0,0,0,0),
    plot.title = element_text(vjust=-0.2)
  )
})

# cluster
pdf(out, width=18, height=12, useDingbats=FALSE)
wrap_plots(plot_list, ncol=6)
dev.off()

# celltype
pdf(out, width=18, height=6.5, useDingbats=FALSE)
wrap_plots(plot_list, ncol=6)
dev.off()

```



RRHO plots to compare DEGs between MHb and LHb
Figures 4C and 5D

```{r eval=FALSE}


library(RRHO)
library(viridis)
library(ggpubr)


degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/DEGs/data/celltype_Nurr2c_vs_GFP.csv')
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype'


cur_target_type <- 'primary'
cur_target_type <- 'secondary'
cur_target_type <- 'other'


degs <- subset(degs, target_type == cur_target_type)


# lowest non-zero value
lowest <- degs %>% subset(p_val_adj != 0) %>% top_n(-1, wt=p_val_adj) %>% .$p_val_adj
degs$p_val_adj <- ifelse(degs$p_val_adj == 0, lowest, degs$p_val_adj)

groups <- c('MHb-Neuron','LHb-Neuron', 'PHb-Neuron')
pairs <- combn(groups, 2)

# plot settings
rrho_plot_list <- list()
cor_list <- c()
NR4A2 <- TRUE
#rrho_maxval <- 500
rrho_maxval <- 350

colfunc <- inferno

for(i in 1:ncol(pairs)){

  cur_pair <- pairs[,i]
  name <- paste(cur_pair, collapse='_')
  print(name)
  cur_x <- cur_pair[1];  cur_y <- cur_pair[2]

  cur_degs_x <- subset(degs, group == cur_x & Nr4a2 == NR4A2 )
  cur_degs_y <- subset(degs, group == cur_y & Nr4a2 == NR4A2 )

  # make sure they are in the same order:
  rownames(cur_degs_x) <- cur_degs_x$gene
  rownames(cur_degs_y) <- cur_degs_y$gene
  cur_degs_y <- cur_degs_y[cur_degs_x$gene,]

  # join the two dataframes
  plot_df <- dplyr::inner_join(cur_degs_x, cur_degs_y, by = 'gene')

  cur_cor <-  cor(x=as.numeric(plot_df$avg_log2FC.x), y=as.numeric(plot_df$avg_log2FC.y))
cor_list <- c(cor_list, cur_cor)

  # set up gene lists
  gl1 <- plot_df[,c('gene', 'avg_log2FC.x')]
  gl2 <- plot_df[,c('gene', 'avg_log2FC.y')]

  # run rrho
  test <- RRHO(gl1, gl2, alternative='enrichment', BY=TRUE)
  overlap_df <- reshape2::melt(test$hypermat.by)
  #subset(overlap_df, value != Inf) %>% .$value %>% max
  overlap_df$value <- ifelse(overlap_df$value > rrho_maxval, rrho_maxval, overlap_df$value)

  # plot rrho heatmap
  p <- ggplot(overlap_df, aes(x=Var1, y=Var2, fill=value, color=value)) +
    ggrastr::rasterise(geom_tile(), dpi=500) +
    scale_fill_gradientn(colors=colfunc(256), limits=c(0, rrho_maxval)) +
    scale_color_gradientn(colors=colfunc(256), limits=c(0, rrho_maxval)) +
    theme(
      plot.title=element_text(hjust=0.5, size=5, face='plain'),
      axis.line=element_blank(),
      axis.ticks=element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.title = element_blank(),
      plot.margin=margin(0,0,0,0)
    ) + coord_equal() + ggtitle(paste0(name, '\nR=', signif(as.numeric(cur_cor),3)))

  rrho_plot_list[[name]] <- p

}


pdf(paste0(fig_dir, 'deg_rrho_combined_', cur_target_type, '.pdf'), width=3, height=7)
wrap_plots(rrho_plot_list, ncol=1) + plot_layout(guides='collect')
dev.off()

```


Quantify overlap between the Naive & Behavior DEGs  
Euler Diagrams in Figures 4 and 5

```{r eval=FALSE}

library(GeneOverlap)

# cluster DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_identity)
name <- 'cluster'
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_identity)
)
degs_nurr2c <- degs

# Naive cluster DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/cluster_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_identity)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_identity)
)
Nr4a2 <- TRUE
degs_naive <- degs

# celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype'
degs_nurr2c <- degs

# Naive celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
Nr4a2 <- FALSE
degs_naive <- degs


#--------------------------------------------------------------
# Compute gene overlaps
#--------------------------------------------------------------

groups <- levels(degs_nurr2c$group)

fc_cutoff <- 0.25
genome.size <- nrow(seurat_obj)
overlap_df <- do.call(rbind, lapply(groups, function(cur_group){

  cur_nurr2c_up <- degs_nurr2c %>% subset(group == cur_group & p_val_adj < 0.05 & avg_log2FC >= fc_cutoff) %>% .$gene
  cur_naive_up <- degs_naive %>% subset(group == cur_group & p_val_adj < 0.05 & avg_log2FC >= fc_cutoff) %>% .$gene
  #up_setsize <- length(intersect)

  cur_nurr2c_down <- degs_nurr2c %>% subset(group == cur_group & p_val_adj < 0.05 & avg_log2FC <= -1* fc_cutoff) %>% .$gene
  cur_naive_down <- degs_naive %>% subset(group == cur_group & p_val_adj < 0.05 & avg_log2FC <= -1* fc_cutoff) %>% .$gene

  cur_overlap_up <- testGeneOverlap(newGeneOverlap(
      cur_nurr2c_up,
      cur_naive_up,
      genome.size=genome.size
  ))
  cur_overlap_down <- testGeneOverlap(newGeneOverlap(
      cur_nurr2c_down,
      cur_naive_down,
      genome.size=genome.size
  ))

  cur_overlap <- data.frame(
    'odds.ratio' = c(cur_overlap_up@odds.ratio, cur_overlap_down@odds.ratio),
    'pval' = c(cur_overlap_up@pval, cur_overlap_down@pval),
    'Jaccard' = c(cur_overlap_up@Jaccard, cur_overlap_down@Jaccard),
    'size_intersection' = c(length(cur_overlap_up@intersection), length(cur_overlap_down@intersection)),
    'group' = c(cur_group, cur_group),
    'direction' = c('up', 'down'),
    'size_set1' = c(length(cur_nurr2c_up), length(cur_nurr2c_down)),
    'size_set2' = c(length(cur_naive_up), length(cur_naive_down)),
    'size_union' = c(length(cur_overlap_up@union), length(cur_overlap_down@union))
  )

  cur_overlap
})) %>% as.data.frame
overlap_df$fdr <- p.adjust(overlap_df$pval)
overlap_df$group <- factor(
  as.character(overlap_df$group),
  levels = groups
)

subset(overlap_df, group == 'LHb-Neuron')
subset(overlap_df, group == 'MHb-Neuron')


#####################################################################
# Euler diagram to compare behavior & naive
#####################################################################

subset(degs_nurr2c, group == 'MHb-Neuron' & avg_log2FC < -0.25 & p_val_adj < 0.05) %>% nrow
subset(degs_naive, group == 'MHb-Neuron' & avg_log2FC < -0.25 & p_val_adj < 0.05) %>% nrow

library(eulerr)

cur_celltype <- 'MHb-Neuron'
cur_celltype <- 'LHb-Neuron'

cur_direction <- 'up'
cur_direction <- 'down'

cur_df <- overlap_df %>% subset(direction == cur_direction & group == cur_celltype)
overlap_list <- c(
  'Behavior' = cur_df$size_set1 - cur_df$size_intersection ,
  'Naive' = cur_df$size_set2 - cur_df$size_intersection ,
  'Behavior&Naive' =  cur_df$size_intersection 
)

pdf(paste0(fig_dir, 'eulerr_', cur_celltype,'_', cur_direction,'.pdf'), width=4, height=4)
plot(euler(overlap_list), quantities = TRUE)
dev.off()


#####################################################################
# Euler diagram to compare behavior & naive
#####################################################################

fc_thresh <- 0.25

degs <- degs_naive; name <- 'naive'
degs <- degs_nurr2c; name <- 'nurr2c'

ct1 <- 'MHb-Neuron'
ct2 <- 'LHb-Neuron'


cur_degs1_up <- subset(degs, group == ct1 & p_val_adj < 0.05 & avg_log2FC > fc_thresh) %>% .$gene
cur_degs1_down <- subset(degs, group == ct1 & p_val_adj < 0.05 & avg_log2FC < -fc_thresh) %>% .$gene

cur_degs2_up <- subset(degs, group == ct2 & p_val_adj < 0.05 & avg_log2FC > fc_thresh) %>% .$gene
cur_degs2_down <- subset(degs, group == ct2 & p_val_adj < 0.05 & avg_log2FC < -fc_thresh) %>% .$gene

cur_overlap_up <- testGeneOverlap(newGeneOverlap(
    cur_degs1_up,
    cur_degs2_up,
    genome.size=genome.size
))
cur_overlap_down <- testGeneOverlap(newGeneOverlap(
    cur_degs1_down,
    cur_degs2_down,
    genome.size=genome.size
))

cur_overlap_up 
cur_overlap_up@odds.ratio
cur_overlap_down
cur_overlap_down@odds.ratio


overlap_list <- c(
  'ct1' = length(setdiff(cur_degs1_up, cur_degs2_up)),
  'ct2' = length(setdiff(cur_degs2_up, cur_degs1_up)),
  'ct1&ct2' =  length(intersect(cur_degs1_up, cur_degs2_up))
)

pdf(paste0(fig_dir, 'eulerr_', name, '_up.pdf'), width=4, height=4)
plot(euler(overlap_list), quantities = TRUE)
dev.off()

overlap_list <- c(
  'ct1' = length(setdiff(cur_degs1_down, cur_degs2_down)),
  'ct2' = length(setdiff(cur_degs2_down, cur_degs1_down)),
  'ct1&ct2' =  length(intersect(cur_degs1_down, cur_degs2_down))
)

pdf(paste0(fig_dir, 'eulerr_', name, '_down.pdf'), width=4, height=4)
plot(euler(overlap_list), quantities = TRUE)
dev.off()


```


Comparison of DEGs between Experienced & Naive Groups

(Figures 4E, 5H-J)

```{r eval=FALSE}

setwd('../tf_net')

# define the sets of primary, secondary, and other genes for Nr4a2

tf_nets <- dir(paste0(data_dir, 'tf_nets/'))
tf_nets <- tf_nets[grepl('importance', tf_nets)]
#cur_net_file <- tf_nets[8]

cur_celltype <- 'MHb-Neuron'
cur_group <- 'Nurr2c'
cur_net_file <- paste0('TFnet_', cur_celltype, '_', cur_group, '_importance.csv')

cur_tf <- 'Nr4a2'

#tf_nets <- tf_nets[grepl(cur_celltype, tf_nets)]

# parameters for regulons
n_tfs <- 5
importance_thresh <- 0.001
combined_output <- data.frame()

primary_genes <- list()
secondary_genes <- list()

for(cur_net_file in tf_nets){

  print(cur_net_file)

  # load the tf-gene table
  importance_df <- read.csv(paste0(data_dir, 'tf_nets/', cur_net_file ))

  tmp <- strsplit(cur_net_file, '_')[[1]]
  cur_celltype <- tmp[2]
  cur_group <- tmp[3]

  #---------------------------------------------------------------------------#
  # Define the TF regulons
  #---------------------------------------------------------------------------#

  regulons <- importance_df %>% 
    subset(Gain > importance_thresh) %>% 
    group_by(gene) %>%
    slice_max(order_by=Gain, n=n_tfs) %>% 
    ungroup()

  # compute the degree for each TF:
  tf_degrees <- table(regulons$tf)

  #---------------------------------------------------------------------------#
  # Get the primary & secondary targets of Nr4a2
  #---------------------------------------------------------------------------#

  # primary target genes 
  cur_primary<- regulons %>% 
    subset(tf == cur_tf) 

  # which of these pimary target genes are tfs?
  cur_primary_tfs <- cur_primary %>% 
    subset(gene %in% unique(regulons$tf)) %>% .$gene

  cur_tfs <- unique(c(cur_tf, cur_primary_tfs))

  # get the regulons for these TFs:
  cur_secondary <- subset(regulons, tf %in% cur_primary_tfs)
  cur_secondary_tfs <- cur_primary %>% 
    subset(gene %in% unique(regulons$tf)) %>% .$gene

  primary_genes[[paste0(cur_celltype, '_', cur_group)]] <- cur_primary$gene 
  secondary_genes[[paste0(cur_celltype, '_', cur_group)]] <- cur_secondary$gene

}



#------------------------------------------------------------------
# Bar plot showing the number of DEGs of each tf target type
#------------------------------------------------------------------

fc_cutoff <- 0.25
barplot_groups <- c('MHb-Neuron', 'LHb-Neuron')
cur_group <- barplot_groups[1]

conditions <- c('Nurr2c', 'GFP'); degs <- degs_nurr2c; name <- 'Behavior'
conditions <- c("NN", 'NGFP'); degs <- degs_naive; name <- 'Naive'

plot_df <- data.frame()
for(cur_group in barplot_groups){

  cur_up <- degs %>% subset(group == cur_group & p_val_adj < 0.05 & avg_log2FC >= fc_cutoff) %>% .$gene
  cur_down <- degs %>% subset(group == cur_group & p_val_adj < 0.05 & avg_log2FC <= -1* fc_cutoff) %>% .$gene

  cur_primary <- unique(unlist(primary_genes[paste0(cur_group, '_', conditions)]))
  cur_secondary <- unique(unlist(secondary_genes[paste0(cur_group, '_', conditions)]))
  cur_secondary <- setdiff(cur_secondary, cur_primary)

  cur_primary_up <- intersect(cur_up, cur_primary)
  cur_primary_down <- intersect(cur_down, cur_primary)
  cur_secondary_up <- intersect(cur_up, cur_secondary)
  cur_secondary_down <- intersect(cur_down, cur_secondary)
  cur_other_up <- setdiff(cur_up, unique(c(cur_primary_up, cur_secondary_up)))
  cur_other_down <- setdiff(cur_down, unique(c(cur_primary_down, cur_secondary_down)))

  cur_df <- data.frame(
    group = cur_group,
    target_type = c('primary', 'primary', 'secondary', 'secondary', 'other', 'other'),
    direction = c('up', 'down', 'up', 'down', 'up', 'down'),
    n = c(
      length(cur_primary_up),
      length(cur_primary_down),
      length(cur_secondary_up),
      length(cur_secondary_down),
      length(cur_other_up),
      length(cur_other_down)
    )
  )

  plot_df <- rbind(plot_df, cur_df)

}

plot_df$target_type <- factor(as.character(plot_df$target_type), levels=c('other', 'secondary', 'primary'))
plot_df$n <- ifelse(plot_df$direction == 'down', -1 * plot_df$n, plot_df$n)
plot_max <- plot_df %>% group_by(direction, group) %>% summarise(x=sum(n)) %>% .$x %>% abs %>% max

p <- plot_df %>%
  ggplot(aes(x = n, y=group, fill=target_type)) + 
  geom_bar(position='stack', stat='identity') + 
  geom_vline(xintercept=0) +
  geom_text(aes(label=abs(n)), position = position_stack(vjust=0.5)) +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank()
  ) + 
  xlab(bquote("N"[genes])) 

pdf(paste0(fig_dir, 'degs_bar_', name,'.pdf'), width=5, height=1.75, useDingbats=FALSE)
p
dev.off()


#---------------------------------------------------------------------------#
# Comparison of DEG effect sizes from experienced and naive groups
#---------------------------------------------------------------------------#

# celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs_nurr2c <- degs

# Naive celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
name <- 'celltype'
degs_naive <- degs

# only keep neurons
degs_nurr2c <- subset(degs_nurr2c, group %in% c('MHb-Neuron', 'LHb-Neuron', 'PHb-Neuron'))
degs_naive <- subset(degs_naive, group %in% c('MHb-Neuron', 'LHb-Neuron', 'PHb-Neuron'))

gene_option <- 'primary'
gene_option <- 'secondary'
gene_option <- 'other'

# plot settings
groups <- unique(degs_nurr2c$group)
plot_list <- list()
cor_list <- c()
signif_only <- FALSE

logfc_thresh <- 0.25

for(cur_group in groups){
  print(cur_group)

  cur_degs1 <- subset(degs_nurr2c, group == cur_group)
  cur_degs2 <- subset(degs_naive, group == cur_group)

  # get the primary / secondary / target genes:
  cur_primary <- unique(unlist(primary_genes[grepl(cur_group, names(primary_genes))]))
  cur_secondary <- unique(unlist(secondary_genes[grepl(cur_group, names(secondary_genes))]))
  if(gene_option == 'primary'){
    tf_genes <-  cur_primary
  } else if (gene_option == 'secondary'){
    tf_genes <- setdiff(cur_secondary, cur_primary)
  } else if (gene_option == 'other'){
    tf_genes <- setdiff(rownames(seurat_obj), unique(c(cur_primary, cur_secondary)))
  }

  genes.keep <- intersect(cur_degs1$gene, cur_degs2$gene)
  genes.keep <- genes.keep[genes.keep %in% tf_genes]
  cur_degs1 <- subset(cur_degs1, gene %in% genes.keep)
  cur_degs2 <- subset(cur_degs2, gene %in% genes.keep)

  # make sure they are in the same order:
  rownames(cur_degs1) <- cur_degs1$gene
  rownames(cur_degs2) <- cur_degs2$gene
  cur_degs2 <- cur_degs2[cur_degs1$gene,]

  tmp1 <- subset(cur_degs1, p_val_adj < 0.05 & avg_log2FC > 0)
  tmp2 <- subset(cur_degs2, p_val_adj < 0.05 & avg_log2FC > 0)

  # join the two dataframes
  plot_df <- dplyr::inner_join(cur_degs1, cur_degs2, by = 'gene')

  subset(plot_df, avg_log2FC.x < -0.25 & avg_log2FC.y < -0.25 &)

  plot_df$group <- ifelse(plot_df$avg_log2FC.x >= logfc_thresh & plot_df$avg_log2FC.y >= logfc_thresh, "Consistent", "")
  plot_df$group <- ifelse(plot_df$avg_log2FC.x <= -logfc_thresh & plot_df$avg_log2FC.y <= -logfc_thresh, "Consistent", plot_df$group)
  plot_df$group <- ifelse(plot_df$avg_log2FC.x >= logfc_thresh & plot_df$avg_log2FC.y <= -logfc_thresh, "Inconsistent", plot_df$group)
  plot_df$group <- ifelse(plot_df$avg_log2FC.x <= -logfc_thresh & plot_df$avg_log2FC.y >= logfc_thresh, "Inconsistent", plot_df$group)
  group_colors <- c('grey', 'dodgerblue', 'seagreen')
  names(group_colors) <- c('', 'Consistent', 'Inconsistent')

  # color the not-significant ones grey
  plot_df$group <- ifelse(plot_df$p_val_adj.y > 0.05 | plot_df$p_val_adj.x > 0.05, "", plot_df$group )

  cur_cor <-  cor(x=as.numeric(plot_df$avg_log2FC.x), y=as.numeric(plot_df$avg_log2FC.y))
  cor_list <- c(cor_list, cur_cor)
  print(table(plot_df$group))
  print(dim(plot_df))

  # how many overlapping?
  up_right <- plot_df %>% subset(avg_log2FC.x >= logfc_thresh & avg_log2FC.y >= logfc_thresh & c(p_val_adj.x < 0.05 & p_val_adj.y < 0.05)) %>% nrow
  down_right <- plot_df %>% subset(avg_log2FC.x >= logfc_thresh & avg_log2FC.y <= -logfc_thresh & c(p_val_adj.x < 0.05 & p_val_adj.y < 0.05)) %>% nrow
  up_left <- plot_df %>% subset(avg_log2FC.x <= -logfc_thresh & avg_log2FC.y >= logfc_thresh & c(p_val_adj.x < 0.05 & p_val_adj.y < 0.05)) %>% nrow
  down_left <- plot_df %>% subset(avg_log2FC.x <= -logfc_thresh & avg_log2FC.y <= -logfc_thresh & c(p_val_adj.x < 0.05 & p_val_adj.y < 0.05)) %>% nrow

annotations <- data.frame(
        xpos = c(-Inf,-Inf,Inf,Inf),
        ypos =  c(-Inf, Inf,-Inf,Inf),
        annotateText = c(as.character(down_left),as.character(up_left), as.character(down_right),as.character(up_right)),
        hjustvar = c(-1,-1,2,2),
        vjustvar = c(-1,2,-1,2)) #<- adjust


  plot_range <- max(abs(plot_df$avg_log2FC.x), abs(plot_df$avg_log2FC.y))
  p <- plot_df %>%
    ggplot(aes(x = avg_log2FC.x, y = avg_log2FC.y, color=group)) +
    annotate("segment", x =logfc_thresh, y = logfc_thresh, yend=Inf, xend=logfc_thresh, linetype='dashed', color='dodgerblue', alpha=0.75) +
    annotate("segment", x =logfc_thresh, y = logfc_thresh, xend=Inf, yend=logfc_thresh, linetype='dashed', color='dodgerblue', alpha=0.75) +
    annotate("segment", x =-logfc_thresh, y = -logfc_thresh, yend=-Inf, xend=-logfc_thresh, linetype='dashed', color='dodgerblue', alpha=0.75) +
    annotate("segment", x =-logfc_thresh, y = -logfc_thresh, xend=-Inf, yend=-logfc_thresh, linetype='dashed', color='dodgerblue', alpha=0.75) +
    annotate("segment", x =-logfc_thresh, y = logfc_thresh, yend=Inf, xend=-logfc_thresh, linetype='dashed', color='seagreen', alpha=0.75) +
    annotate("segment", x =-logfc_thresh, y = logfc_thresh, xend=-Inf, yend=logfc_thresh, linetype='dashed', color='seagreen', alpha=0.75) +
    annotate("segment", x =logfc_thresh, y = -logfc_thresh, yend=-Inf, xend=logfc_thresh, linetype='dashed', color='seagreen', alpha=0.75) +
    annotate("segment", x =logfc_thresh, y = -logfc_thresh, xend=Inf, yend=-logfc_thresh, linetype='dashed', color='seagreen', alpha=0.75) +
    geom_vline(xintercept =0, linetype='dashed', color='lightgrey') +
    geom_hline(yintercept =0, linetype='dashed', color='lightgrey') +
    ggrastr::rasterise(geom_point(), dpi=500) +
    scale_color_manual(values=group_colors) +
   # geom_smooth(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC.x, y = avg_log2FC.y), method='lm', color='black') +
   # stat_cor(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC.x, y = avg_log2FC.y), method='pearson') +
    xlim(c(-plot_range, plot_range)) +
    ylim(c(-plot_range, plot_range)) +
    theme(
      axis.line.x = element_blank(),
      axis.line.y = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      plot.title = element_text(hjust=0.5)
    ) +
    coord_fixed(ratio=1) + NoLegend() +
    xlab(bquote("Behavior Avg. log"[2]~"(FC)")) +
    ylab(bquote("Naive Avg. log"[2]~"(FC)")) +
    ggtitle(cur_group)  +
    geom_text(inherit.aes=FALSE, data=annotations,aes(x=xpos,y=ypos,hjust=hjustvar,vjust=vjustvar,label=annotateText))

  plot_list[[cur_group]] <- p

}


out <- paste0(fig_dir, 'deg_corr_combined_',gene_option,'.pdf')

plot_list <- lapply(plot_list, function(x){
  x + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(0,0,0,0),
    plot.title = element_text(vjust=-0.2)
  )
})


# cell_type
pdf(out, width=9, height=6)
wrap_plots(plot_list, ncol=3)
dev.off()

pdf(out, width=18, height=6.5)
wrap_plots(plot_list, ncol=6)
dev.off()



```

GO term enrichment analysis for DEGs 

Figures 4F-H and 5F-G

```{r eval=FALSE}

library(enrichr)

dbs<-c('GO_Biological_Process_2018','GO_Cellular_Component_2018','GO_Molecular_Function_2018', 'WikiPathways_2019_Mouse', 'KEGG_2019_Mouse')

dir.create(paste0(fig_dir, name, '/enrichr/'))

# celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype'
Nr4a2 <- FALSE
nlabel <- 5
color1 <- 'darkgoldenrod3'; color2 <- 'hotpink3'

# Naive celltype DEGs
degs <- read.csv(file='/dfs7/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/revision/DEGs/data/celltype_Naive_Nurr2c_vs_GFP_DEGs.csv')
degs <- subset(degs, group %in% seurat_obj$cell_type)
degs$group <- factor(
  degs$group,
  levels = levels(seurat_obj$cell_type)
)
name <- 'celltype_naive'
Nr4a2 <- TRUE
nlabel <- 5
color2 <- "#643D78"; color1 <- "#14703F"

# subset the neurons (this doesn't work for cluster level info)
degs <- subset(degs, group %in% c('MHb-Neuron', 'LHb-Neuron', 'PHb-Neuron'))
degs$group <- droplevels(degs$group)

clusters <- unique(degs$group)

tmp <- data.frame()
for(cur_celltype in clusters){



combined_output <- data.frame()
for(cur_group in clusters){

  print(cur_group)
  genes.up <- degs %>% subset(group == cur_group & p_val_adj <= 0.05 & avg_log2FC >= 0.25)  %>% .$gene %>% as.character
  genes.down <- degs %>% subset(group == cur_group & p_val_adj <= 0.05 & avg_log2FC <= -0.25)  %>% .$gene %>% as.character

  enriched_up <- enrichr(genes.up,dbs)
  enriched_down <- enrichr(genes.down,dbs)

  for(db in dbs){
    cur_df_up <- enriched_up[[db]]
    cur_df_down <- enriched_down[[db]]

    if (nrow(cur_df_up) > 1){
      cur_df_up$db <- db
      cur_df_up$group <- cur_group
      cur_df_up$upregulated <- TRUE
      combined_output <- rbind(combined_output, cur_df_up)
    }

    if (nrow(cur_df_down) > 1){
      cur_df_down$db <- db
      cur_df_down$group <- cur_group
      cur_df_down$upregulated <- FALSE
      combined_output <- rbind(combined_output, cur_df_down)
    }

  }
}

write.csv(combined_output, file=paste0('enrichr_tables/', name, '_GO_terms.csv'), quote=FALSE, row.names=FALSE)

#--------------------------------------------------------#
# Plotting
#--------------------------------------------------------#

# re-load table:
combined_output <- read.delim(paste0(data_dir, 'Nurr2c_Neuron_DEGs_GO_terms_signif.tsv'), sep='\t', header=1)

selected_terms <- read.delim(paste0(data_dir, 'Nurr2c_Neuron_DEGs_GO_terms_selected.txt'), sep='\t', header=1)
selected_terms$Term <- str_replace(selected_terms$Term, " \\s*\\([^\\)]+\\)", "")

p <- selected_terms  %>% subset(group == 'MHb-Neuron') %>%
  ggplot(aes(x=-log(Combined.Score), y=reorder(Term, Combined.Score), fill=group))+
  geom_bar(stat='identity', position='identity') +
  geom_text(aes(label=Term), x=-.1, color='black', size=3.5, hjust='right') +
  scale_fill_manual(values=celltype_colors) +
  xlab('log(Enrichment)') +
  scale_x_continuous(expand = c(0, 0), limits = c(NA, 0)) +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    legend.title = element_blank(),
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    axis.line.y=element_blank(),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_blank()
  )

pdf(paste0(fig_dir, 'Nurr2c_selected_GO_terms_down_bar.pdf'), width= 3, height=3 , useDingbats=FALSE)
p + facet_wrap(~group, ncol=1, scales='free') + NoLegend()
dev.off()


p <- selected_terms  %>% subset(group == 'LHb-Neuron') %>%
  ggplot(aes(x=log(Combined.Score), y=reorder(Term, Combined.Score), fill=group))+
  geom_bar(stat='identity', position='identity') +
  geom_text(aes(label=Term), x=.1, color='black', size=3.5, hjust='left') +
  scale_fill_manual(values=celltype_colors) +
  xlab('log(Enrichment)') +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    legend.title = element_blank(),
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    axis.line.y=element_blank(),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_blank()
  )

pdf(paste0(fig_dir, 'Nurr2c_selected_GO_terms_up_bar.pdf'), width= 3, height=2 , useDingbats=FALSE)
p + facet_wrap(~group, ncol=1, scales='free') + NoLegend()
dev.off()

```

