

In this notebook we take .bam files from cell ranger, and the fully processed
Seurat object, to make a bigwig track for each cell group that we can vive
on the UCSC genome browser.

# 1: Load Data

```{r eval=FALSE}
# conda activate spatial

library(Seurat)
library(tidyverse)

# load seurat obj:
seurat_obj <- readRDS('/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/data/harmony_annotated.rds')

load('/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/data/color_scheme.rda')

setwd("/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs")

```

# Step 1: Get Barcode lists for each cell type in each sample
For each sample, make a .tsv file that tells sinto how to split up the .bams

```{r eval=FALSE}

################################################################################
# split by celltype
################################################################################

dir.create('data/barcodes/')

# first combine COP with OPC:
seurat_obj$cell_type <- ifelse(as.character(seurat_obj$cell_type) == 'COP', 'OPC', as.character(seurat_obj$cell_type))

samples <- unique(seurat_obj$Sample) %>% as.character

# loop over each sample
for(sample in samples){

  # print cell barcodes for one whole Sample:
  barcodes <- seurat_obj@meta.data %>%
    subset(Sample == sample)

  # add cell type column
  out_df <- data.frame(
    'barcode'=paste0(do.call(rbind, strsplit(rownames(barcodes), '-'))[,1], '-1'),
    'file' = gsub(' ', '_', barcodes$cell_type)
  )

  name <- gsub("Sample", "Wood_Lab", sample)
  name <- gsub('-', '_', name)
  print(name)

  # write
  write.table(out_df, file=paste0('data/barcodes/', name, '_celltype_barcodes.tsv'), sep='\t', col.names=FALSE, row.names=FALSE, quote=FALSE)

}


################################################################################
# split by cluster
################################################################################

dir.create('data/barcodes/')

samples <- unique(seurat_obj$Sample) %>% as.character

# loop over each sample
for(sample in samples){

  # print cell barcodes for one whole Sample:
  barcodes <- seurat_obj@meta.data %>%
    subset(Sample == sample)

  # add cell type column
  out_df <- data.frame(
    'barcode'=paste0(do.call(rbind, strsplit(rownames(barcodes), '-'))[,1], '-1'),
    'file' = gsub(' ', '_', barcodes$annotation)
  )

  name <- gsub("Sample", "Wood_Lab", sample)
  name <- gsub('-', '_', name)
  print(name)

  # write
  write.table(out_df, file=paste0('data/barcodes/', name, '_cluster_barcodes.tsv'), sep='\t', col.names=FALSE, row.names=FALSE, quote=FALSE)

}


```

# Step 2: Run `sinto_split-bams.sub` script

```{r eval=FALSE}

cd ~/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/analysis/trackhubs/run_sinto

# run the script!
sbatch ../bin/sinto_split-bams.sub

```


## Step 3: write list of files to combined using samtools

```{r eval=FALSE}


################################################################################
# cell types
################################################################################

bam_dir <- '/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/data/bams/'

groups <- seurat_obj$cell_type  %>% as.character %>% unique
groups <- gsub(' ', '_', groups)

# Nurr2C
samples <- seurat_obj@meta.data %>% subset(Group == 'Nurr2c') %>% .$Sample %>% unique %>% as.character
samples <- gsub("Sample", "Wood_Lab", samples)
samples <- gsub('-', '_', samples)

for(group in groups){
  print(group)
  bam_files <- paste0(bam_dir, samples, '_celltype/', group, '.bam')
  bam_files <- bam_files[file.exists(bam_files)]
  write.table(bam_files, file=paste0('data/bam_merge_lists/', group, '_Nurr2c.tsv'), sep='\t', col.names=FALSE, row.names=FALSE, quote=FALSE)
}


# GFP
samples <- seurat_obj@meta.data %>% subset(Group == 'GFP') %>% .$Sample %>% unique %>% as.character
samples <- gsub("Sample", "Wood_Lab", samples)
samples <- gsub('-', '_', samples)
for(group in groups){
  bam_files <- paste0(bam_dir, samples, '_celltype/', group, '.bam')
  bam_files <- bam_files[file.exists(bam_files)]
  write.table(bam_files, file=paste0('data/bam_merge_lists/', group, '_GFP.tsv'), sep='\t', col.names=FALSE, row.names=FALSE, quote=FALSE)
}


################################################################################
# clusters
################################################################################

bam_dir <- '/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/data/bams_clusters/'

groups <- seurat_obj$annotation  %>% as.character %>% unique
groups <- gsub(' ', '_', groups)

# Nurr2C
samples <- seurat_obj@meta.data %>% subset(Group == 'Nurr2c') %>% .$Sample %>% unique %>% as.character
samples <- gsub("Sample", "Wood_Lab", samples)
samples <- gsub('-', '_', samples)

for(group in groups){
  print(group)
  bam_files <- paste0(bam_dir, samples, '_cluster/', group, '.bam')
  bam_files <- bam_files[file.exists(bam_files)]
  write.table(bam_files, file=paste0('data/bam_merge_lists/', group, '_Nurr2c.tsv'), sep='\t', col.names=FALSE, row.names=FALSE, quote=FALSE)
}


# GFP
samples <- seurat_obj@meta.data %>% subset(Group == 'GFP') %>% .$Sample %>% unique %>% as.character
samples <- gsub("Sample", "Wood_Lab", samples)
samples <- gsub('-', '_', samples)
for(group in groups){
  bam_files <- paste0(bam_dir, samples, '_cluster/', group, '.bam')
  bam_files <- bam_files[file.exists(bam_files)]
  write.table(bam_files, file=paste0('data/bam_merge_lists/', group, '_GFP.tsv'), sep='\t', col.names=FALSE, row.names=FALSE, quote=FALSE)
}


```

## Step 4: Merge bam files based on lists we made in step 3:

```{r eval=FALSE}

cd ~/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/run_merge
sbatch ../bin/merge_bams.sub

```

## Step 5: Convert Merged .bams to bigwig format:

```{bash eval=FALSE}

cd ~/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/run_bam_to_bigwig
sbatch ../bin/bam_to_bigwig.sub


# re-name .bw to .bigWig
cd ~/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/data/bigwigs/

for file in ./*.bw
do
  name=$(basename $file .bw)
  mv $file $name.bigWig
done


```




## Step 6: Make necessary files for UCSC browser:

* genomes.txt file
* trackDb.txt file
* hub.txt file


track ODC_i_AD
bigDataUrl ODC_i_AD.bigWig
shortLabel ODC_i_AD
longLabel ODC_i_AD
type bigWig
visibility full
color 255,171,255
autoScale on

```{r eval=FALSE}

colors <- c(celltype_colors, cluster_colors)

bigwig_dir = "/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/data/bigwigs/"
bigwigs <- dir(bigwig_dir)

scale_factor <- 1.6

outfile <- "/dfs3b/swaruplab/smorabit/collab/woodlab/cocaine_mouse_2021/Nurr2c_vs_GFP/trackhubs/data/trackDb.txt"

for(bw in bigwigs){

  # get name
  cur_name <- strsplit(bw, '[.]')[[1]][1]
  cur_anno <- strsplit(cur_name, '[_]')[[1]][1]

  # get color
  cur_color <- colors[strsplit(bw, '[_]')[[1]][1]]
  cur_rgb <- grDevices::col2rgb(cur_color) %>% as.numeric %>% paste0(collapse=",")

  # get group
  cur_group <- strsplit(bw, '[_]')[[1]][2]
  cur_group <- strsplit(cur_group, '[.]')[[1]][1]

  # how many cells in this group?
  if(grepl("Neuron", cur_name)){
    cur_nCells <- subset(seurat_obj@meta.data, cell_type == cur_anno & Group == cur_group) %>% nrow
  } else{
    cur_nCells <- subset(seurat_obj@meta.data, annotation == cur_anno & Group == cur_group) %>% nrow
  }

  # scale number of cells for visualiztion
  cur_nCells <- round(cur_nCells * scale_factor)
  cur_view <- paste0("0:", cur_nCells)
  print(cur_view)

  # write all info
  lines <- list(
    paste("track", cur_name),
    paste("bigDataUrl", bw),
    paste("shortLabel", cur_name),
    paste('longLabel', cur_name),
    paste('type bigWig'),
    paste('visibility full'),
    paste('color', cur_rgb),
    paste('autoScale on'),
  #  paste('viewLimits', cur_view),
    paste("")
  )

  for(l in lines){
    cat(l, file = outfile, sep="\n", append=TRUE)
  }

}





```

## Step 7: Moving stuff to the right spot

* Move all bigwigs to artemis
* put them in a web-facing folder such as

/var/www/html/Nurr2c_trackhubs/

* point the UCSC browser to this folder and voila!

Put them here for this paper:
https://swaruplab.bio.uci.edu/Nurr2c_trackhubs/
